---
title: "Hawaii Plant METE Results Log"
author: "A. J. Rominger"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

## function to color code deviations from theory, also useful for state variable plotting
fitCol <- function(x) {
    m <- 1/(max(x, na.rm = TRUE) - min(x, na.rm = TRUE))
    
    newx <- m * (x - min(x, na.rm = TRUE))
    
    colval <- colorRamp(hsv(c(0.17, 0.6), c(0.5, 1), c(1, 0.7)))(newx)
    colval[is.na(colval)] <- 0
    
    col <- rgb(colval, maxColorValue = 255)
    col[col == '#000000'] <- 'transparent'
    
    return(col)
}
```

Setup workspace and load data
```{r}
library(socorro)
library(meteR)
library(sp)
setwd('~/Dropbox/Research/hawaiiMETE_plant')

load('meteSumm.RData')
load('meteEg.RData')
load('~/Dropbox/hawaiiDimensions/kokua/data/islands.RData')
```

First look at how deviations from METE play out across the islands. Fit is based on $z^2$ values as discussed in Rominger and Merow MEE paper.
```{r}
layout(matrix(2:1, nrow = 2), heights = c(1, 4))
par(mar = rep(0, 4))
plot(meteSumm, col = fitCol(sqrt(meteSumm@data$z2SAD)), pch = 16, cex = 1)
plot(spTransform(islands, CRS(proj4string(meteSumm))), add = TRUE)

par(mar = c(0, 3, 0, 0) + 0.5, mgp = c(2, 0.75, 0))
plot(sort(meteSumm@data$z2SAD), col = fitCol(sort(sqrt(meteSumm@data$z2SAD))), 
     pch = 16, cex = 1, ylab = 'z^2 value', xaxt = 'n')
```

Overall the distribution of $z^2$ values tells us METE is pretty much never rejected (critical value at $\alpha = 0.05$ is $\chi_{1, P=0.95}^2 =$ `r round(qchisq(0.95, 1), 3)`.  We can look at a few (random) example SADs:

```{r}
par(mfrow = c(4, 4), mar = c(2, 2, 0, 0), oma = c(2, 2, 0.5, 0.5), 
    mgp = c(1.5, 0.3, 0), tcl = -0.25)
for(i in 1:length(meteEg)) {
    plot(meteEg[[i]]$sad, ptype = 'rad', add.legend = FALSE, 
         xlab = '', ylab = '', log = 'y', ylim = c(1, max(meteEg[[i]]$sad$data)), 
         yaxt = 'n')
    logAxis(2)
}
mtext('Rank', side = 1, outer = TRUE)
mtext('Abundance', side = 2, outer = TRUE)
```

When sample size is small (i.e. a few species) it's hard to make concrete conclusions about fit. The $z^2$ value is based on simulations that hold $S_0$ and $N_0$ fixed, so should hypothetically account for sample size.

We can look explicitly at fit across variables of interest
```{r}
par(mfrow = c(1, 4), mar = c(3, 1, 0, 0) + 0.5, oma = c(0, 2, 0, 0), mgp = c(2, 0.75, 0))
plot(meteSumm@data[, c('age_mid', 'z2SAD')], log = 'x', xaxt = 'n'); logAxis(1)
plot(meteSumm@data[, c('MAP', 'z2SAD')])
plot(meteSumm@data[, c('MAT', 'z2SAD')])
plot(meteSumm@data[, c('Prop_Introduced', 'z2SAD')])
mtext('z^2 value', side = 2, outer = TRUE, line = 0.5)
```


Looking now jointly at environment and age as predictors of deviation
```{r}
layout(matrix(2:1, nrow = 2), heights = c(1, 4))
par(mar = c(3, 3, 0, 0) + 0.5, mgp = c(2, 0.75, 0))
plot(meteSumm@data[, c('age_mid', 'MAP')], log = 'x', xaxt = 'n', pch = 16, cex = 1, 
     col = fitCol(sqrt(meteSumm@data$z2SAD)))
logAxis(1)
par(mar = c(0, 3, 0, 0) + 0.5, mgp = c(2, 0.75, 0))
plot(sort(meteSumm@data$z2SAD), col = fitCol(sort(sqrt(meteSumm@data$z2SAD))), 
     pch = 16, cex = 1, ylab = 'z^2 value', xaxt = 'n')
```

Checking if state variables influence fit, they don't
```{r}
par(mfrow = c(1, 3), mar = c(3, 1, 0, 0) + 0.5, oma = c(0, 2, 0, 0), mgp = c(2, 0.75, 0))
plot(meteSumm@data[, c('S0', 'z2SAD')])
plot(meteSumm@data[, c('N0', 'z2SAD')], log = 'x', xaxt = 'n'); logAxis(1)
plot(meteSumm@data[, c('E0', 'z2SAD')], log = 'x', xaxt = 'n'); logAxis(1)
mtext('z^2 value', side = 2, outer = TRUE, line = 0.5)
```

Speaking of state variables, how are they correlated?
```{r}
pairs(meteSumm@data[, c('S0', 'N0', 'E0')], log = 'xy')
```

How do state variables change with age and plot area?
```{r}
par(mfcol = c(4, 2), mar = c(2, 2, 1, 1), mgp = c(2, 0.75, 0), oma = c(1, 1, 0, 0) + 0.5)
plot(meteSumm@data[, c('max_PlotArea', 'z2SAD')], log = 'x')
mtext('z^2 value', side = 2, line = 2)

plot(meteSumm@data[, c('max_PlotArea', 'S0')], log = 'xy')
mtext('S0', side = 2, line = 2)

plot(meteSumm@data[, c('max_PlotArea', 'N0')], log = 'xy')
mtext('N0', side = 2, line = 2)

plot(meteSumm@data[, c('max_PlotArea', 'E0')], log = 'xy')
mtext('E0', side = 2, line = 2)
mtext('Plot area', side = 1, line = 2)

plot(meteSumm@data[, c('age_mid', 'z2SAD')], log = 'x')
plot(meteSumm@data[, c('age_mid', 'S0')], log = 'xy')
plot(meteSumm@data[, c('age_mid', 'N0')], log = 'xy')
plot(meteSumm@data[, c('age_mid', 'E0')], log = 'xy')
mtext('Age', side = 1, line = 2)
```

```{r}
layout(matrix(2:1, nrow = 2), heights = c(1, 4))
par(mar = c(3, 3, 0, 0) + 0.5, mgp = c(2, 0.75, 0))
plot(meteSumm@data[, c('age_mid', 'max_PlotArea')], log = 'xy', pch = 16, cex = 1, 
     col = fitCol(sqrt(meteSumm@data$z2SAD)))
par(mar = c(0, 3, 0, 0) + 0.5, mgp = c(2, 0.75, 0))
plot(sort(meteSumm@data$z2SAD), col = fitCol(sort(sqrt(meteSumm@data$z2SAD))), 
     pch = 16, cex = 1, ylab = 'z^2 value', xaxt = 'n')
```

We can also look for age-specific SARs
```{r}
layout(matrix(2:1, nrow = 2), heights = c(1, 4))
par(mar = c(3, 3, 0, 0) + 0.5, mgp = c(2, 0.75, 0))
plot(meteSumm@data[, c('age_mid', 'max_PlotArea')], log = 'xy', pch = 16, cex = 1, 
     col = fitCol(log(meteSumm@data$S0)))
par(mar = c(0, 3, 0, 0) + 0.5, mgp = c(2, 0.75, 0))
plot(sort(meteSumm@data$S0), col = fitCol(sort(log(meteSumm@data$S0))), 
     pch = 16, cex = 1, ylab = 'S_0', xaxt = 'n', log = 'y')
```


